### 2018-06-13 papers/synth/intro.tex
asd
fixed typos

%Motivate VTS
Eukaryotic cells, including human cells, consist of multiple membrane
bound compartments.
%
Material is transported among these compartments by the vesicle
transport system (VTS).
%
Briefly, the source compartment produces a membrane bound packet of
molecules called a vesicle.
%
After release, this vesicle specifically recognizes the correct target
compartment within the cell, and fuses with it.
%
A lot of information about the molecules that form the machinery of
the VTS has been discovered, including their regulatory interaction
with each other. 
% %
% \todo{connect to function }{We also know a lot about the specific functions} of these molecules and how these molecules interact amongst each other.
%
In spite of this detailed knowledge at the level of the molecules, the
structure of the VTS network, or the road-map of the eukaryotic cell,
is far from complete.
%
The current knowledge of the network is put together from a patch work
of biological experiments and is scattered across several publications.
%
Even after this information is collected and put together, we find
that the network obtained is still not complete; new vesicles and new
contents in previously known vesicles are constantly being discovered.
%
The synthesis for the unknown pieces may be assisted by computation on
the graph model of VTSs.
%
% 
%In this work, we use a model described in [] to find out, given a partial
%VTS graph and details about interactions amongst its molecules, what the complete graph of the
%VTS could look like. (Please look at appendix A for the detailed discussion on pros and cons of the
%model.
%
%Eukaryotic cells, including human cells, consists of multiple compartments.
%
%Vesicle Traffic Systems(VTSs) are the material transport mechanisms among the compartments inside the cells~\cite{vtsIntro}.
%
%Almost all subsystems of the cells depend on VTSs.
%
%Therefore, understanding how VTSs functions is one of the key
%questions of cell biology.
%
In this paper, we are looking at the computational questions 
arise from the VTSs.

We use the model of VTSs that has been presented in~\cite{shukla2017discovering}.
%
Please look at appendix~\ref{sec:model}
for the detailed discussion on pros and cons of the model.
% Describe VTS in detail

VTSs are regulated by the same molecules that they transport. Molecules on compartments and on vesicles can be in either an active or an inactive state.
% the term vesicle has already been introduced.. using two terms for the same object seems confusing. 
%
% (The molecules are transported to the other compartments via
%unidirectional channels.) cannot comment on unidirectionality of vesicles. they may be undergoingmultiple cycles of back-fusion for all we %know..  
%
The active molecules regulate the vesicle they occur on.
%
A vesicle is enabled by a set of molecules
such that one part of the set occurs in the vesicle and the other part
occurs in the target compartment.
% use either target or destination consistantly throughout..i prefer target because that's what the t in t-SNARE stands for.
%
These molecules must be active in the vesicle
and target compartment respectively.
%
The pairs are called {\em fusing} sets and analogously
the vesicle is considered to be {\em fused} with the
destination compartment.
%
Not all sets molecule can participate in the fusion, the
potential fusion pairs are usually determined by biological experiments.
%
Generally, in cells, the fusing pairs are found to be distinct for distinct vesicle-compartment fusions.
%
In order that a molecule that has participated in a fusion does not interfere
with fusion at compartments and vesicles, in the model, we require that the molecule is inactive on appropriate compartments.
%
%
The activity of molecules is regulated by the other molecules, i.e.,
the presence and absence of the other molecules in a compartment or
vesicle may make the molecule active or inactive.
%
We call this regulation as {\em activity functions}.

@@

% Partial info defence
%
As we have discussed earlier, our understanding of VTSs is partial.
%
% It largely means that the current knowledge is due to the patch work
% of biological experiments~\cite{model} for identifying the compartments,
% channels, molecules, and their activity and regulatory control.
% %
% Often the information is scattered around in several publications.
% %
% Even after with some effort we may collate all the known information about
% a VTS, we may still have unknown pieces of the system; new vesicles and new contents in previously known vesicles are constantly being discovered.
%
%{Speculative: Make a case for completing partial parts of VTSs!!}
%
%For example, it is possible that some edges may not have been
%discovered in a VTS that does not satisfy the stability condition.
% nature of missing information has already been elaborated upon in the first paragraph
%
The synthesis for the unknown pieces may be {\em assisted} by computation on
the graph model of VTSs.
%
In this paper, we consider several versions of the synthesis problem
involving different parts of VTSs that can be synthesized, such as
modifying labels, adding/deleting edges, and learning activity functions.
%
We also consider variation on the properties against which we do
synthesis, namely stability, and $k$-connectedness that states that
VTS remains connected after removing any $k-1$ edges.
%
In order to synthesize the parts of a VTS such that it satisfies the
constraints, we encode the synthesis problem into a satisfiability of
quantified Boolean formulas(QBFs). 
%

We have implemented the encoding in a flexible tool,
which can handle a wide range of synthesis query.
%
We have applied our tool on several VTSs including two found-in-nature
VTSs.
%

% \input{fig-mukund-vts}
% %
% In figure~\ref{fig:mukund-vts}, we present one of the two for
% mammalian cells obtained by studying the
% literature.\footnote{Please read appendix~\ref{sec:ex-vts} for more biological details.}
% %
% The VTS has six nodes and 55 molecules.
% %
% We can easily check that stability condition is not satisfied for
% many molecules for example ???.
% %
% On the example, our synthesis tool reports that minimum ?? edges needed
% to be added to make the VTS stable.
% %
% The dashed edges are the suggested edges from our tool.
% %

>>>>

%Motivate VTS
Eukaryotic cells, including human cells, consist of multiple membrane
bound compartments.
%
Material is transported among these compartments by the vesicle
transport system (VTS).
%
Briefly, the source compartment produces a membrane bound packet of
molecules called a vesicle.
%
After release, this vesicle specifically recognizes the correct target
compartment within the cell, and fuses with it.
%
A lot of information about the molecules that form the machinery of
the VTS has been discovered, including their regulatory interaction
with each other. 
% %
% \todo{connect to function }{We also know a lot about the specific functions} of these molecules and how these molecules interact amongst each other.
%
In spite of this detailed knowledge at the level of the molecules, the
structure of the VTS network, or the road-map of the eukaryotic cell,
is far from complete.
%
The current knowledge of the network is put together from a patch work
of biological experiments and is scattered across several publications.
%
Even after this information is collected and put together, we find
that the network obtained is still not complete; new vesicles and new
contents in previously known vesicles are constantly being discovered.
%
The synthesis for the unknown pieces may be assisted by computation on
the graph model of VTSs.
%
% 
%In this work, we use a model described in [] to find out, given a partial
%VTS graph and details about interactions amongst its molecules, what the complete graph of the
%VTS could look like. (Please look at appendix A for the detailed discussion on pros and cons of the
%model.
%
%Eukaryotic cells, including human cells, consists of multiple compartments.
%
%Vesicle Traffic Systems(VTSs) are the material transport mechanisms among the compartments inside the cells~\cite{vtsIntro}.
%
%Almost all subsystems of the cells depend on VTSs.
%
%Therefore, understanding how VTSs functions is one of the key
%questions of cell biology.
%
In this paper, we are looking at the computational questions 
arising from the VTSs.

We use the model of VTSs that has been presented in~\cite{shukla2017discovering}.
%
Please look at appendix~\ref{sec:model}
for a detailed discussion on pros and cons of the model.
% Describe VTS in detail

VTSs are regulated by the same molecules that they transport. Molecules on compartments and on vesicles can be in either an active or an inactive state.
% the term vesicle has already been introduced.. using two terms for the same object seems confusing. 
%
% (The molecules are transported to the other compartments via
%unidirectional channels.) cannot comment on unidirectionality of vesicles. they may be undergoingmultiple cycles of back-fusion for all we %know..  
%
The active molecules regulate the vesicle they occur on.
%
A vesicle is enabled by a set of molecules
such that one part of the set occurs in the vesicle and the other part
occurs in the target compartment.
% use either target or destination consistantly throughout..i prefer target because that's what the t in t-SNARE stands for.
%
These molecules must be active in the vesicle
and target compartment respectively.
%
The pairs are called {\em fusing} sets and analogously
the vesicle is considered to be {\em fused} with the
destination compartment.
%
Not all molecule sets can participate in the fusion, the
potential fusion pairs are usually determined by biological experiments.
%
Generally, in cells, the fusing pairs are found to be distinct for distinct vesicle-compartment fusions.
%
In order that a molecule that has participated in a fusion does not interfere
with fusion at compartments and vesicles, in the model, we require that the molecule is inactive on appropriate compartments.
%
%
The activity of molecules is regulated by the other molecules, i.e.,
the presence and absence of the other molecules in a compartment or
vesicle may make the molecule active or inactive.
%
We call this regulation as {\em activity functions}.

@@

% Partial info defence
%
As we have discussed earlier, our understanding of VTSs is partial.
%
% It largely means that the current knowledge is due to the patch work
% of biological experiments~\cite{model} for identifying the compartments,
% channels, molecules, and their activity and regulatory control.
% %
% Often the information is scattered around in several publications.
% %
% Even after with some effort we may collate all the known information about
% a VTS, we may still have unknown pieces of the system; new vesicles and new contents in previously known vesicles are constantly being discovered.
%
%{Speculative: Make a case for completing partial parts of VTSs!!}
%
%For example, it is possible that some edges may not have been
%discovered in a VTS that does not satisfy the stability condition.
% nature of missing information has already been elaborated upon in the first paragraph
%
The synthesis for the unknown pieces may be {\em assisted} by computation on
the graph model of VTSs.
%
In this paper, we consider several versions of the synthesis problem
involving different parts of VTSs that can be synthesized, such as
modifying labels, adding/deleting edges, and learning activity functions.
%
We also consider variation on the properties against which we do
synthesis, namely stability, and $k$-connectedness that states that the VTS remains connected after removing any $k-1$ edges.
%
In order to synthesize the parts of a VTS such that it satisfies the
constraints, we encode the synthesis problem into one of satisfiability of
quantified Boolean formulas(QBFs). 
%

We have implemented the encoding in a flexible tool,
which can handle a wide range of synthesis queries.
%
We have applied our tool on several VTSs including two found-in-nature
VTSs.
%
% \input{fig-mukund-vts}
% %
% In figure~\ref{fig:mukund-vts}, we present one of the two for
% mammalian cells obtained by studying the
% literature.\footnote{Please read appendix~\ref{sec:ex-vts} for more biological details.}
% %
% The VTS has six nodes and 55 molecules.
% %
% We can easily check that stability condition is not satisfied for
% many molecules for example ???.
% %
% On the example, our synthesis tool reports that minimum ?? edges needed
% to be added to make the VTS stable.
% %
% The dashed edges are the suggested edges from our tool.
% %

### 2018-06-13 papers/synth/prelim.tex
add
fixed typos

\begin{df}
  A VTS $G$ is a tuple $(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$, where
  \begin{itemize}
  \item $\nodes$ is a set of nodes representing compartments in the VTS,
  \item $\mols$ is the set of molecules flowing in the system, 
  \item $\edges \subseteq \nodes \times (\powerset{\mols}-\emptyset) \times \nodes$ is the
    set of edges with molecule sets as labels,
  \item $\nlabel : \nodes \maps \powerset{\mols}$ defines the molecules present in the nodes,
  \item $\pairs \subseteq 2^{\mols}$ is pairing relation,
  \item $\nodef : \mols \maps \powerset{\mols} \maps \booleans$ is activity maps for nodes, and
  \item $\edgef : \mols \maps \powerset{\mols} \maps \booleans $ is activity maps for edges.
  \end{itemize}
\end{df}
$\nodes$, $\mols$, $\edges$, and $\nlabel$ define a labelled graph.
%
Additionally, $\pairs$ defines which molecules can fuse with which molecules,
and
$\nodef$ and $\edgef$ are the activity functions for molecules on
nodes and edges respectively.
%
The model captures the study state of a VTS.
%
The analysis of the model will inform us about the network/graph
properties of VTSs.

@@

The following constraints encode the basic consistency of VTSs.
\begin{align*}
  \texttt{EdgeC} =\;&\hspace{-1ex}\bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
  \land
  \bigwedge\limits_{i,q} \neg e_{i,i,q}
  \land
  % \tag{\texttt{NodeC}}\label{eq:f0}\\
  \bigwedge\limits_{\mathclap{i,j,q,m}} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
  \\
  \texttt{ActivityC} =\;&
  \bigwedge\limits_{\mathclap{i,j,q,m}} b_{i,j,q,m} \limplies e_{i,j,q,m} \quad\land\quad
  \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
  % \tag{V2}\label{eq:f1}\\
  \\
  \texttt{PairingC} =\;&
  \exists qr. \Land\limits_{\mathrlap{m_1,m_2,m_3,m_4}\;\;}
                         (p_{\{m_1,m_2,m_3,m_4\}} \limplies qr_{m_1} + qr_{m_2} + qr_{m_3} + qr_{m_4} = 3) 
% \land\\ 
%   & \quad\quad\quad ( m_1 = m_2 = m_3 = m_4  \limplies \lnot p_{m_1,m_2,m_3,m_4} )
  % \bigwedge\limits_{m} 
  % (\Lor\limits_{m'} p_{mm'} \limplies \Land_{m'} \lnot p_{m'm} ) \land
  % (\Lor\limits_{m'} p_{m'm} \limplies \Land_{m'} \lnot p_{mm'} )
  \\
  \texttt{Fusion1} =\;&
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
  \bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
                        \Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
  & \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
  \\
  \texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
  & \hspace{2cm} \neg 
  \bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
  \\
  \texttt{Consistancy} =\;& \texttt{EdgeC} \land
  \texttt{ActivityC} \land \texttt{PairingC} \land
  \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
\texttt{EdgeC} states that each edge has at least one molecule,
there are no self loops, and edge labels are consistent with node labels.   
\texttt{ActivityC} states that active molecule are present.
\texttt{PairingC} states that all molecules are divided into
two types and any fusing set of molecules
must have three molecules involved from one type and one molecule from the other.
\texttt{Fusion1}, and \texttt{Fusion2} states the well-fused condition.
\texttt{Consistancy} is the conjunction of all of the above.

@@

For synthesis of incomplete systems,
we need properties  against which we synthesize the missing parts.
%
Here we will discussed two such properties proposed in earlier
works~\cite{smtVTS}, namely stability and $k$-connectedness.
%

>>>>

\begin{df}
  A VTS $G$ is a tuple $(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$, where
  \begin{itemize}
  \item $\nodes$ is a set of nodes representing compartments in the VTS,
  \item $\mols$ is the set of molecules flowing in the system, 
  \item $\edges \subseteq \nodes \times (\powerset{\mols}-\emptyset) \times \nodes$ is the
    set of edges with molecule sets as labels,
  \item $\nlabel : \nodes \maps \powerset{\mols}$ defines the molecules present in the nodes,
  \item $\pairs \subseteq 2^{\mols}$ is pairing relation,
  \item $\nodef : \mols \maps \powerset{\mols} \maps \booleans$ is activity maps for nodes, and
  \item $\edgef : \mols \maps \powerset{\mols} \maps \booleans $ is activity maps for edges.
  \end{itemize}
\end{df}
$\nodes$, $\mols$, $\edges$, and $\nlabel$ define a labelled graph.
%
Additionally, $\pairs$ defines which molecules can fuse with which molecules,
and
$\nodef$ and $\edgef$ are the activity functions for molecules on
nodes and edges respectively.
%
The model captures the steady state of a VTS.
%
The analysis of the model will inform us about the network/graph
properties of VTSs.

@@

The following constraints encode the basic consistancy of VTSs.
\begin{align*}
  \texttt{EdgeC} =\;&\hspace{-1ex}\bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
  \land
  \bigwedge\limits_{i,q} \neg e_{i,i,q}
  \land
  % \tag{\texttt{NodeC}}\label{eq:f0}\\
  \bigwedge\limits_{\mathclap{i,j,q,m}} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
  \\
  \texttt{ActivityC} =\;&
  \bigwedge\limits_{\mathclap{i,j,q,m}} b_{i,j,q,m} \limplies e_{i,j,q,m} \quad\land\quad
  \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
  % \tag{V2}\label{eq:f1}\\
  \\
  \texttt{PairingC} =\;&
  \exists qr. \Land\limits_{\mathrlap{m_1,m_2,m_3,m_4}\;\;}
                         (p_{\{m_1,m_2,m_3,m_4\}} \limplies qr_{m_1} + qr_{m_2} + qr_{m_3} + qr_{m_4} = 3) 
% \land\\ 
%   & \quad\quad\quad ( m_1 = m_2 = m_3 = m_4  \limplies \lnot p_{m_1,m_2,m_3,m_4} )
  % \bigwedge\limits_{m} 
  % (\Lor\limits_{m'} p_{mm'} \limplies \Land_{m'} \lnot p_{m'm} ) \land
  % (\Lor\limits_{m'} p_{m'm} \limplies \Land_{m'} \lnot p_{mm'} )
  \\
  \texttt{Fusion1} =\;&
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
  \bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
                        \Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
  & \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
  \\
  \texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
  & \hspace{2cm} \neg 
  \bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
  \\
  \texttt{Consistancy} =\;& \texttt{EdgeC} \land
  \texttt{ActivityC} \land \texttt{PairingC} \land
  \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
\texttt{EdgeC} states that each edge has at least one molecule,
there are no self loops, and edge labels are consistent with node labels.   
\texttt{ActivityC} states that active molecule are present.
\texttt{PairingC} states that all molecules are divided into
two types and any fusing set of molecules
must have three molecules involved from one type and one molecule from the other.
\texttt{Fusion1}, and \texttt{Fusion2} states the well-fused condition.
\texttt{Consistancy} is the conjunction of all of the above.

@@

For synthesis of incomplete systems,
we need properties  against which we synthesize the missing parts.
%
Here we will discuss two such properties proposed in earlier
works~\cite{smtVTS}, namely stability and $k$-connectedness.
%
